---
title: "Week 5 Application Part"
output: html_document
---

## The cut Function
Converts a quantitative variable to a grouping factor. 
For instance, you may wish to present your data in terms of
age in 5-year groups.

install.packages("ISwR")
```{r echo=TRUE}
library(ISwR)
attach(juul)
head(juul)
summary(juul)
age <- subset(juul, age >= 10 & age <= 16)$age
range(age)
agegr <- cut(age, seq(10,16,2))
length(age)
table(agegr)
agegr2 <- cut(age, seq(10,16,2), right=F)
table(agegr2)
```

## The quantile Function
```{r echo=TRUE}
q <- quantile(age, c(0, .25, .50, .75, 1))
q
ageQ <- cut(age, q)
table(ageQ)
```

## Separating data frames
Let's separate juul data set for boys and girls.
The net effect of the negative indices is to remove, 
for example, testvol and sex from juulgrl.
```{r echo=TRUE}
juulgrl <- subset(juul, sex==2, select=-c(testvol,sex))
juulboy <- subset(juul, sex==1, select=-c(menarche,sex))
head(juulgrl)
head(juulboy)
```

## Merging data frames
Merge ? adds variables to a dataset. 
Merging two datasets require that both have at least one variable in common (either string or numeric). If string make sure the categories have the same spelling (i.e. country names, etc.). 

Two data frames are called respectively dfx and dfy. The merged frame is computed as merge(dfx, dfy, by="ID").
```{r echo=TRUE}
mrg <- merge(juulboy, juulgrl, by=c("age","tanner","igf1"))
head(mrg,10)
```

## Getting the Titanic Data
```{r echo=TRUE}
train.data<-read.csv(file="https://raw.githubusercontent.com/agconti/kaggle-titanic/master/data/train.csv",header=T,sep=",")
str(train.data)
```

```{r echo=TRUE}
train.data$Pclass = factor(train.data$Pclass)
train.data$Survived = factor(train.data$Survived)
str(train.data)
```

```{r echo=TRUE}
is.na(train.data$Age)
sum(is.na(train.data$Age) == TRUE)
length(train.data$Age)
sum(is.na(train.data$Name)==T)
length(train.data$Name)
```

# For a title containing a missing value, assign the mean age value for each title not containing a missing value.

```{r echo=TRUE}
mean.mr = mean(train.data$Age[grepl(" Mr\\.",
     train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Mr\\.",train.data$Name) & is.na(train.data$Age)]=mean.mr
```

```{r echo=TRUE}
mean.mrs = mean(train.data$Age[grepl(" Mrs\\.", 
      train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Mrs\\.",train.data$Name)&is.na(train.data$Age)]=mean.mrs
```
```{r echo=TRUE}
mean.dr = mean(train.data$Age[grepl(" Dr\\.",
      train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Dr\\.",train.data$Name) & is.na(train.data$Age)]=mean.dr
```
```{r echo=TRUE}
mean.miss = mean(train.data$Age[grepl(" Miss\\.", 
      train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Miss\\.",train.data$Name) & is.na(train.data$Age)]=mean.miss
```
```{r echo=TRUE}
mean.master = mean(train.data$Age[grepl(" Master\\.",
       train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Master\\.", train.data$Name) & is.na(train.data$Age)]=mean.master
```

# List the distribution of Port of Embarkation. 
Add the useNA = "always" argument to show the number of NA values
```{r echo=TRUE}
table(train.data$Embarked, useNA = "always")
```

# Assign the two missing values to the most counted port, which is Southampton in this case:
```{r echo=TRUE}
train.data$Embarked[which(is.na(train.data$Embarked))] = 'S'
table(train.data$Embarked, useNA = "always")
```

# Exploring and visualizing data
```{r echo=TRUE}
barplot(table(train.data$Survived), main="Passenger Survival",names= c("Perished", "Survived"))
```
```{r echo=TRUE}
barplot(table(train.data$Pclass), main="Passenger Class", names= c("first","second", "third"))
```
```{r echo=TRUE}
barplot(table(train.data$Sex), main="Passenger Gender")
hist(train.data$Age, main="Passenger Age", xlab = "Age")
```
```{r echo=TRUE}
counts = table( train.data$Survived,train.data$Sex)
barplot(counts, col=c("darkblue","red"),
             legend = c("Perished",  "Survived"), 
               main = "Passenger Survival by Sex")
```
```{r echo=TRUE}
hist(train.data$Age[which(train.data$Survived == "0")], main= "Passenger Age Histogram", xlab="Age", ylab="Count", col ="blue", breaks=seq(0,80,by=2))
hist(train.data$Age[which(train.data$Survived == "1")], col ="red", add=T, breaks=seq(0,80,by=2))
```
